---
title: "Prediction Assignment Writeup"
output: html_document
---

Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

"Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E)."

This human activity recognition research has traditionally focused on discriminating between different activities, i.e. to predict "which" activity was performed at a specific point in time (like with the Daily Living Activities dataset above).   

In this project I will try to predict "which" activity (exactly according to the specication (Class A) , throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).) was performed at a specific point in time.   


The data is downloaded from the http://groupware.les.inf.puc-rio.br/har website. The caret package will be used in the prediction.   

## Loading and preprocessing the data

```{r loaddata}
    library(caret)
    training <- read.csv("pml-training.csv", head = T)
    testing <- read.csv("pml-testing.csv",, head = T)

    # check if there are missing value in the trai and test datasets
    sum(is.na(training))
    sum(is.na(testing))
```

## Preprocessing the data (remove missing variables)
There are quite a few variables in the pml-training and pml-testing datasets with missing values. These variables are dropped from the two datasets.   

```{r preparation}
training <- training[,c("user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", 
                "new_window", "num_window", "roll_belt", "pitch_belt", "yaw_belt", "total_accel_belt", 
                "gyros_belt_x", "gyros_belt_y", "gyros_belt_z", "accel_belt_x", "accel_belt_y", 
                "accel_belt_z", "magnet_belt_x", "magnet_belt_y", "magnet_belt_z", "roll_arm", 
                "pitch_arm", "yaw_arm", "total_accel_arm", "gyros_arm_x", "gyros_arm_y", 
                "gyros_arm_z", "accel_arm_x", "accel_arm_y", "accel_arm_z", "magnet_arm_x", 
                "magnet_arm_y", "magnet_arm_z", "roll_dumbbell", "pitch_dumbbell", 
                "yaw_dumbbell", "gyros_dumbbell_x", "gyros_dumbbell_y", "gyros_dumbbell_z", 
                "accel_dumbbell_x", "accel_dumbbell_y", "accel_dumbbell_z", "magnet_dumbbell_x",
                "magnet_dumbbell_y", "magnet_dumbbell_z", "roll_forearm", "pitch_forearm", 
                "yaw_forearm", "total_accel_forearm", "gyros_forearm_x", "gyros_forearm_y", 
                "gyros_forearm_z", "accel_forearm_x", "accel_forearm_y", "accel_forearm_z", 
                "magnet_forearm_x", "magnet_forearm_y", "magnet_forearm_z", "classe")]

testing <- testing[,c("user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", 
                        "new_window", "num_window", "roll_belt", "pitch_belt", "yaw_belt", "total_accel_belt", 
                        "gyros_belt_x", "gyros_belt_y", "gyros_belt_z", "accel_belt_x", "accel_belt_y", 
                        "accel_belt_z", "magnet_belt_x", "magnet_belt_y", "magnet_belt_z", "roll_arm", 
                        "pitch_arm", "yaw_arm", "total_accel_arm", "gyros_arm_x", "gyros_arm_y", 
                        "gyros_arm_z", "accel_arm_x", "accel_arm_y", "accel_arm_z", "magnet_arm_x", 
                        "magnet_arm_y", "magnet_arm_z", "roll_dumbbell", "pitch_dumbbell", 
                        "yaw_dumbbell", "gyros_dumbbell_x", "gyros_dumbbell_y", "gyros_dumbbell_z", 
                        "accel_dumbbell_x", "accel_dumbbell_y", "accel_dumbbell_z", "magnet_dumbbell_x",
                        "magnet_dumbbell_y", "magnet_dumbbell_z", "roll_forearm", "pitch_forearm", 
                        "yaw_forearm", "total_accel_forearm", "gyros_forearm_x", "gyros_forearm_y", 
                        "gyros_forearm_z", "accel_forearm_x", "accel_forearm_y", "accel_forearm_z", 
                        "magnet_forearm_x", "magnet_forearm_y", "magnet_forearm_z", "problem_id")]

```

## Mode Fitting
I use randomforest algorithm to predict "which" activity (exactly according to the specication (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E)) was performed at a specific point in time. The prediction is performed on testing data.   

1) k-fold Cross Validation
The k-fold cross validation method involves splitting the dataset into k-subsets. For each subset is held out while the model is trained on all other subsets. This process is completed until accuracy is determine for each instance in the dataset, and an overall accuracy estimate is provided.

In this prediction, 10-fold cross validation method is used as follows:   
```{r corss-validation}
    # define training control
    train_control <- trainControl(method="cv", number=10)
```

2) Fit a model
Randome Forest algorithm model is used. Model fit from the train function, I use the 19622 samples in a training set and 57 predictors. The model uses corss-validation with 10 fold.   
```{r modelfit}
    modelFit <- train(classe ~., data=training, trControl=train_control, method="rf")
```
## Estimating Model Accuracy
Accuracy estimate shows the model is %99.1 accurate.   

```{r mode accuracy}
    modelFit
```

## Prediction new values on testing data
I predict the model on the test dataset as follows:   
```{r prediction}
    predictions <- predict(modelFit, newdata=testing)
    # Here are set of predications 
    predictions
```